---
title: "data_exclusions"
author: 'Selin Bekir'
date: 'March 16, 2022'
output: html_notebook
---

#load packages
```{r}
require(dplyr)
require(ggplot2)
require(tidyr)
require(stringr)
```

# load all-finished-runs.csv file
## this csv file is exported from Cognition.run. Only the SONA IDs and IP addresses were deleted, before sharing the data public
```{r}
df <- read.csv('/divided_visual_field_go-no-go_task/Data/all-finished.csv')
glimpse(df)
```

### EXCLUSIONS ###
## BLOCKS ##
# accuracy
1. excluding all <.50 blocks
```{r}
#creating a separate data.frame for 
df_acc_block_counts <- df %>% group_by(run_id, cond_order, accuracy == 1) %>% mutate(acc_block_counts = n())
df_acc_block_counts <- filter(df_acc_block_counts, accuracy == 1)
df_acc_block_counts <- subset(df_acc_block_counts, select = c(run_id, cond_order, acc_block_counts))
df_acc_block_counts <- df_acc_block_counts[!duplicated(df_acc_block_counts), ]

df <- merge(df, df_acc_block_counts, by=c("run_id", "cond_order"), all=T)

#exclude all acc_count <30/60 blocks:
df <- filter(df, acc_block_counts > 30)
#summary(df_mains$acc_block_counts)
```

2. excluding maximizers in < .75 acc blocks
```{r}
#looking for potential maximizers in <.75 proportion correct blocks (45/60 trials)
df_maxim <- filter(df, acc_block_counts <= 45)
df_maxim <- df_maxim %>% group_by(run_id, cond_order, response) %>% mutate(response_counts = n())
df_maxim <- subset(df_maxim, select = c(run_id, target_shown, main, cond_order, response, response_counts))
df_maxim <- filter(df_maxim, target_shown == 'pink')
df_maxim  <- df_maxim[!duplicated(df_maxim ), ]
df_maxim_respond <- filter(df_maxim, main == 'respond')
df_maxim_respond <- filter(df_maxim_respond, response == 'null')
df_maxim_inhibit <- filter(df_maxim, main == 'inhibit')
df_maxim_inhibit <- filter(df_maxim_inhibit, response == 'k')
df_maximizers <- rbind(df_maxim_respond, df_maxim_inhibit)
#put 1 for maximizers, 0 for none maximzers
df_maximizers <- df_maximizers %>% mutate(maxim = case_when(response_counts >= 57 ~ 1,
                                             response_counts < 57 ~ 0)) 
df_maximizers <- subset(df_maximizers, select = c(run_id, cond_order, maxim ))
df_maximizers <- filter(df_maximizers, maxim == 1)

df <- merge(df, df_maximizers, by=c("run_id", "cond_order" ), all=T)
df$maxim <- as.integer(as.logical(df$maxim))
#only get the NAs i.e. non-maximizers
df <- subset(df, is.na(maxim))
df <- subset(df, select = -c(maxim))
```

#fixation
excluding all blocks with < 55 fixation
```{r}
df$fixation_check <- as.numeric(df$fixation_check)
#summary(df$fixation_check)
df <- filter(df, fixation_check > 55)
summary(df$fixation_check)
```

#excludingn participants who have less than 5 blocks remained
```{r}
df <- df %>% group_by(run_id) %>% mutate(trial_counts = n())
df <- filter(df, trial_counts > 300)
```

checking the remaining sample size and sample size per condition
```{r}
n_distinct(df$run_id)
#sample size per condition
df %>%
  group_by(main) %>%
  summarise(count = n_distinct(run_id))

#total trials per subject
df %>%  group_by(run_id) %>% count(main)
```
##TRIALS##
#rt <= 100 exclusions
```{r}
df <- filter(df, is.na(rt)| rt > 100)
#summary(df$rt)
```


#writing it to a new csv file
```{r}
write.csv(df,"~/all-clean.csv", row.names = FALSE)
```


















