---
title: "Logistic Regression on Accuracy Scores"
author: "Selin Bekir"
date: "March 16, 2022"
output: html_notebook
---

load packages
```{r}
require(aod)
require(ggplot2)
require(effects)
library(dplyr)
require(tidyverse)
require(lme4)
require(ggpubr)
require(rstatix)
require(emmeans)
require(sjmisc)
source('~/theme_Publication.R')
require(sjPlot)
require(sjlabelled)
require(sjmisc)
``` 

```{r}
df <- read.csv("~/divided_visual_field_go-no-go_task/Data/all-clean.csv")

#only getting right-handed people
df <- filter(df, handedness >= 20)

#factoring the labels
df <- df %>% mutate(run_id = factor(run_id)) 
levels(df$main) <- c('inhibit', 'respond')
levels(df$cond) <- c('LVF', 'BOTH', 'RVF')
levels(df$target_shown)  <- c('pink', 'green')
```

# total sample size
```{r}
n_distinct(df$run_id)
```

### Logistic Regression ####
DV: trial-by-trial accuracy scores (0-1)
IV1, visual field condition: cond (BOTH, LVF, RVF), within-subject, fixed variable
IV2, main version of the task: main (inhibit, respond), between-subject, fixed variable
IV3, rare or common targets: target_shown (pink, green), within-subject, fixed variable
IV4, subjects: run_id, subjects, random variable

# Using GLMER #
```{r}
logit_glmer <- glmer(accuracy ~ cond * main * target_shown + (1|run_id), data=df, family=binomial(link=logit))
summary(logit_glmer)
```

# table of estimates with 95% CI
```{r}
se <- sqrt(diag(vcov(logit_glmer)))
(tab <- cbind(Est = fixef(logit_glmer), LL = fixef(logit_glmer) - 1.96 * se, UL = fixef(logit_glmer) + 1.96 * se))
```

#odd ratios
```{r}
odds <- exp(tab)
odds
```

# odds ratio plots
```{r}
plot_model(logit_glmer, vline.color = 'gray', show.values = T, value.offset = .4, rm.terms = c('condLVF:mainrespond:target_shownpink', "condRVF:mainrespond:target_shownpink", "mainrespond:target_shownpink", "condRVF:target_shownpink", "condLVF:target_shownpink ")) +theme_Publication()
```
# probability plots
```{r}
plot_model(logit_glmer, transform = "plogis", show.values = T, value.offset = .4, rm.terms = c('condLVF:mainrespond:target_shownpink', "condRVF:mainrespond:target_shownpink", "mainrespond:target_shownpink"))
```
#overall effect of visual field seems to be significant, but what about the pairise comparisons between BOTH-LVF and RVF?
```{r}
emmeans(logit_glmer, list(pairwise ~ cond), adjust = "tukey")
```

```{r}
plot(allEffects(logit_glmer))
```





